---
title: "USDA Validation Report"
author: "Gage Rowden"
format:
  pdf:
    fig-width: 6.5
    fig-height: 4.5
    fig-align: center
    documentclass: report
    header-includes:
      - \usepackage{fontspec}
      - \usepackage{caption}
      - \setmainfont{Calibri}
      - \AtBeginDocument{\floatplacement{table}{H}}
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    fontsize: 12pt
  html:
    embed-resources: true
    code-background: true
    page-layout: full
format-links: false
editor: source
---
## RAMALT ROC Analysis
RAMALT biopsies taken at time zero were considered the negative control group, and those taken at time of necropsy were considered the positive control group. Three metrics, AUC, MPR, and MS were analyzed to determine the sensitivity and specificity of the two assays and dilution factors.

Data is first loaded from Airtable using the airtabler API wrapper. The library can be installed with `{r} remotes::install_github("https://github.com/bergant/airtabler")`. The user must also have a personal access token with the appropriate permissions to access the data.
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(airtabler)

APP <- "app7KsgYl2jhOnYg7"

# Get the necessary tables
tables <- airtable(APP, c("animals", "results"))

results <- tables$results$select_all(
  filterByFormula = "{sample_type} = 'RAMALT'"
)

animals <- tables$animals$select_all()
animals <- rename(animals, "animal" = "animal_id")
```

First the data was cleaned and formatted. 
```{r warning=FALSE, message=FALSE}
library(janitor)

df_ <- results %>%
  mutate(across(everything(), as.character)) %>%
  left_join(animals, by = "animal") %>%
  clean_names() %>%
  mutate(
    across(c(sample_id, animal), as.factor),
    assay = factor(assay, levels = c("RT-QuIC", "Nano-QuIC")),
    mpi   = as.integer(mpi),
    across(c(mpr, raf, ttt, ms, auc), as.numeric)
  )
```
Next, the data was subset to include only the control groups.
```{r}
df_roc <- df_ %>%
  filter(mpi == 0 | mortem == "post-mortem") %>%
  mutate(
    positive = as.integer(mortem == "post-mortem")
  ) %>%
  pivot_longer(cols=c("mpr", "ms", "auc"))
```
Combinations of the relevant variables (metrics, assay, dilution) were generated for making ROCs of subset data. Additionally, a couple of functions for generating the ROCs and formatting data.
```{r message=FALSE}
library(pROC)

# Build all combinations of variables
combos <- expand.grid(
  m = unique(df_roc$name),
  a = unique(df_roc$assay),
  d = unique(df_roc$dilution),
  stringsAsFactors = FALSE
)

# Function to compute ROC + coords
compute_roc <- function(m, a, d) {
  sub_df <- df_roc %>%
    filter(name == m, assay == a, dilution == d)
  
  sub_roc <- roc(
    sub_df, response = "positive", predictor = "value", quiet=TRUE
  )
  sub_roc$metric   <- m
  sub_roc$assay    <- a
  sub_roc$dilution <- d
  
  coord_df <- coords(sub_roc) %>%
    mutate(metric = m, assay = a, dilution = d)
  
  list(roc = sub_roc, coords = coord_df)
}

# Metric Labeller (capitalizes the metric acronyms)
relabel_metrics <- function(x) {
  lvls = sort(unique(x))
  factor(x, levels = lvls, labels = toupper(lvls))
}
```
### RAMALT ROC Results
Apply the ROC function to the sub-sampled datasets.
```{r message=FALSE}
library(kableExtra)

roc_results  <- pmap(combos, compute_roc)
rocs         <- map(roc_results, "roc")
coord_df     <- map(roc_results, "coords") %>% 
  bind_rows() %>%
  mutate(metric = relabel_metrics(metric))

# AUC summary table
auc_df <- combos %>%
  mutate(
    auc         = map_dbl(rocs, auc),
    specificity = 0.65,
    sensitivity = ifelse(a == "RT-QuIC", 0.15, 0.05),
    label       = paste0(a, " AUC = ", signif(auc, 3)),
    m = relabel_metrics(m)
  ) %>%
  rename(assay = a, dilution = d, metric = m) %>%
  arrange(desc(auc))
kable(auc_df)
```

```{r echo=FALSE}
coord_df %>%
  group_by(assay, dilution, metric) %>%
  arrange(sensitivity) %>%
  ggplot(aes(specificity, sensitivity, color = assay)) +
  geom_step(linewidth = 1) +
  geom_text(
    aes(label = label, color = assay),
    data = auc_df, hjust = 0, size = 2,
    show.legend = FALSE
  ) +
  scale_color_manual(values=c("darkslateblue", "darkorange")) +
  facet_grid(vars(dilution), vars(metric)) +
  scale_x_reverse() +
  labs(y = "Sensitivity", x = "Specificity") +
  theme(
    legend.position = "none",
    legend.title = element_blank()
  )
```



